{\itshape Advanced Robotics} A simple path optimization method for motion planning

### 2015-10-26 paper.tex
Add Florent and JP corrections and new convergence figure

\jvol{30} \jnum{15} \jyear{2016} \jmonth{August}

\title{{\itshape Advanced Robotics} A simple path optimization method for motion planning }

\author{Myl\`{e}ne Campana$^{\ast}$ \thanks{$^\ast$Corresponding author. Email: mcampana@laas.fr}, Florent Lamiraux and Jean-Paul Laumond
\\\vspace{6pt}
{\em{LAAS-CNRS, 7 av. Colonel Roche, 31400 Toulouse, FRANCE}}
}
\maketitle

@@

\section{Introduction}

Motion planning for systems in cluttered environments has been addressed for more
than thirty years~\citep{ref-motionplan},~\citep{ref-pianomover}. Most planners
randomly sample the system configuration space~\cite{ref-spaceconfig} in
order to find a collision-free path, linking the initial and final configurations. 
The main issue using these techniques is that the computed path makes unnecessary 
detours and needs to be post-processed before being executed by a virtual or real 
robot. Alternative strategies exist however.
\begin{itemize}
\item Planning by path-optimization~\cite{itomp2012},~\cite{voronoiOMP} where
obstacle avoidance is handled by constraints or cost using computation of the
nearest obstacle distance. Most of these planners are using non-linear
optimization~\cite{BettsNonlinopt} under constraints. Such planners provide close-to-optimality paths and have smaller time computation
for easy problems, but they are mostly unable to solve narrow passage issues.

\item Optimal random sampling~\cite{KaramanPRMstarRRTstar} are also close to an
optimal solution, but computation time is significantly higher than classical
approaches. Moreover the roadmap is only valid for a given motion planning
   problem. 
\end{itemize}

@@

For this shortening purpose, random shortcut (RS) methods are
still very 
popular~\cite{Sekhavat-Svestka1998,Geraerts04clearancebased,HauserFastSmooth}. 
However, RS requires fine 
tuning of the termination condition (see 
(Algorithm~\ref{algo:random-shortcut})) and is no efficient due to randomness for long 
trajectories where only a small part needs to be optimized. 
(Figure~\ref{decoupled_DOF_optimization}) presents another situation where RS will 
always fail to optimize the initial path, since it cannot decouple the 
robot degrees of freedom (DOF) on which the optimization occurs, contrary to our 
algorithm or partial random shortcut (PRS) methods~\cite{Geraerts04clearancebased} 
that shortcut only one or a group of DOF.

\begin{figure}[H]
	\centering
	\includegraphics[width=7.5cm]{decoupled_DOF_optimization.pdf}
	\caption{Example of a path which random shortcut will never manage to 
	optimize: each shortcut tentative will provide a collision or will not 
	decrease the path length. The optimal path belongs to the $x-y$ plane 
	containing $\conf_{init}$ and $\conf_{final}$.}
	\label{decoupled_DOF_optimization}
\end{figure}

On the other hand, numerical
optimization methods like CHOMP~\cite{RatliffChomp} can be used as a
post-processing step. They have clear termination conditions, but collision
avoidance is handled by inequality constraints sampled at many points along
the trajectory. These methods therefore require a pre-processing step of the
robot (and/or environment) model in order to make 
it simpler:~\cite{RatliffChomp} covers PR2 bodies with spheres, 
while~\cite{convexOptimMotplan} needs to decompose objects into convex subsets.
Moreover, in the applications we address, optimality is not desirable as
such since the shortest path between two configurations in a cluttered
environment usually contains contacts with the obstacle,
or is most of the time
out of the scope of numerical optimization algorithms~\cite{LaumondOptim}.

@@

This paper is based on a similar work sent to a conference, where a different type of collision constraints was studied.

\section{RELATED WORK}
Previous work on path optimization has been conducted. CHOMP algorithm~\cite{RatliffChomp} optimizes an initial guess provided as
input. It minimizes a time invariant cost function using efficient covariant
hamiltonian gradient descent. The cost is quantified by non-smooth parts (with
high velocities) and an obstacle avoidance term, provided by the distance to the 
nearest obstacle for each iteration of the trajectory. Calculating these nearest 
distances however is time-consuming because the distances between all pairs of 
objects must be computed at each time step along the path. To reduce the 
computation time, the method starts by building offline a map of distances that 
will be called during the optimization at the requested time. Besides, meshes 
are pre-processed into bounding spheres so that distances are computed faster 
at the cost of a geometry approximation.

@@

Some heuristics use random shortcuts on the initial guess combined with a 
trajectory re-building that returns ${C}^1$ shortcuts made of parabolas and lines 
(bang-bang control)~\cite{HauserFastSmooth}. These local refined trajectories 
are time-optimal since they comply with acceleration and velocity constraints. 
~\cite{Guernane2011}, also based on random shooting, is guiding configuration-
generation with local holonimic considerations. But this method remains only 
locally optimal, and is not addressing high-DOF problems.
~\cite{Geraerts04clearancebased}, whereas just implemented for freeflyer robots, is 
using medial axis retractation for clearance and a PRS (applied only on random DOF) 
which can address the problem of (Figure~\ref{decoupled_DOF_optimization}) but is 
relatively slower than RS. Moreover, PRS is not using the information returned by 
the checker of which links are colliding to guide which DOF or group of DOF are 
relevant to shortcut, contrary to our method.

In some way, our method shares similarities with~\cite{PanSmoothSplineShort} 
since this latter work relies on collision checking and backtracks when an 
iteration is detected in collision, instead of trying to constantly satisfy 
distance constraints. Unlike our method however, anti-collision constraints are 
handled by interpolating configurations which, at some points of the trajectory, 
freeze the whole robot configuration instead of a relevant subpart. Thus, this 
method cannot solve the (Figure~\ref{decoupled_DOF_optimization}) problem.

@@

A robot is defined by a kinematic chain composed of a tree of joints. We denote 
by $(J_1,\cdots,J_m)$ the ordered list of joints. Each joint $J_i$, $1\leq i\leq 
m$ is represented by a mapping from a sub-manifold of $\real^{n_i}$, where $n_i$ 
is the dimension of $J_i$ in the configuration space, to the space 
of rigid-body motions $SE(3)$. The rigid-body motion is the position of the joint 
in the frame of its parent. In the examples shown in this paper, we consider 4 
types of joints described in (Table~\ref{tab:joints}).
\begin{table}[b]
\centerline {
  \begin{tabular}{cccc}
    Name & dimension & config space & velocity\\
    \hline
    translation & 1 & $\real$ & $\real$\\
    bounded rotation & 1& $\real$ & $\real$\\
    unbounded rotation & 2 & $\Sone\subset\real^2$ & $\real$\\
    $SO(3)$ & 4 & $\Sthree\subset\real^4$ & $\real^3$
  \end{tabular}
}
\caption {Translation and rotation joint position are defined by 1 
parameter corresponding respectively to the translation along an axis and a 
rotation angle around an axis. Unbounded rotation is defined by a point on the 
unit circle of the plane: 2 parameters corresponding to the cosine and the 
sine of the rotation angle. $SO(3)$ is defined by a unit quaternion. The 
velocity of translation and unbounded rotation joints is the derivative of the 
configuration variable. The velocity of an unbounded rotation joint corresponds 
to the angular velocity. The velocity of a $SO(3)$ joint is defined by the 
angular velocity vector $\omega\in\real^3$.}
\label{tab:joints}
\end{table}
A configuration of the robot
$$\conf = (\underbrace{q_1,\cdots,q_{n_1}}_{J_1},\underbrace{q_{n_1+1},
\cdots,q_{n_1+n_2}}_{J_2},\cdots q_n),\ n\triangleq\sum_{i=1}^m n_i$$
is defined by the concatenation of the joint configurations. The configuration 
space of the robot is denoted by $\CS\subset\real^n$.

@@

\paragraph {Operations on configurations and vectors} by analogy with the case 
where the configuration space is a vector space, we define the following 
operators between configurations and vectors:
$$
\conf_2 - \conf_1 \in \real^p, \ \conf_1, \conf_2\in\CS
$$
is the constant velocity moving from $\conf_1$ to $\conf_2$ in unit time, and
$$
\conf + \dot{\conf}\in\CS, \ \conf\in\CS \ \dot{\conf}\in\real^p
$$
is the configuration reached from $\conf$ after following constant velocity $
\dot{\conf}$ during unit time.

@@

\subsubsection {Cost}

let $W\in\real^{p\times p}$ be a diagonal matrix of weights:
$$
W=\left(\begin{array}{cccccccccc}
w_1 I_{p_1}       &        &  0  \\
    & w_2 I_{p_2} &        &   \\
    &            & \ddots &   \\
  0 &            &        & w_m I_{p_m}
\end{array}\right)
$$
where $I_{p_i}$ is the identity matrix of size $p_i$ and $w_i$ is the weight 
associated to the lever arm of the joint $J_i$. We define the length of the straight interpolation 
between two configurations as:
$$
\|\conf_2 - \conf_1\|_{W} \triangleq \sqrt{(\conf_2 - \conf_1)^T W^2 (\conf_2 - 
\conf_1)}.
$$
Weights are used to homogenize translations and rotations in the velocity vector. 
For rotations, the weight is equal to the maximal distance of the robot bodies 
moved by the joint to the center of the joint.

Given $\conf_0$ and $\conf_{wp+1}$ fixed, the cost we want to minimize is defined 
by
$$
\cost (\xx) \triangleq \frac{1}{2}\sum_{i=1}^{wp+1} \lambda_{i-1} \|\conf_{i}-\conf_{i-1}\|_{W}
^{2}
$$
where the $\lambda_{i-1}$ coefficients benefit will be explained in the results 
section. For now we can assume that $\forall i,\, \lambda_{i-1}=1$.

@@

\subsection {Resolution}
We assume that the direct interpolation between the initial and final configurations contains collisions.
Let $H$ denotes the constant Hessian of the cost function, an iteration is 
described as follow:
\begin{equation}\label{eq:iteration-1}
\begin{split}
& \pii =  -H^{-1} \nabla c(\xx_i)^{T} \\
& \xx_{i+1} =  \xx_{i} + \alpha_i \pii
\end{split} 
\end{equation}
where $\alpha_i$ is a real valued parameter. Taking $\alpha_i=1$ yields the 
unconstrained minimal cost path, i.e. all waypoints aligned on the straight line 
between $q_0$ and $q_{n+1}$. Since this solution is in collision, we set 
$\alpha_i = \alpha$ 
where $\alpha$ is a parameter that will be explained in the algorithm section.

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=9cm]{optim_grad.pdf}
	\caption{Illustration of one iteration of our optimization. $\xx_{i+1}$ 
	appears 
	to be in collision with the obstacle, the first colliding configuration 
	$\xx_{i+1}(\tcolli)$ is returned by the continuous collision checker. The 
	corresponding constraint will be computed in the backtracked configuration 
	$\xx_{i}(\tcolli)$.}
	\label{optim_grad}
\end{figure}

@@

\subsection*{New constraint}

Let $T$ be a positive real number such that each path $\xx_i$ is a mapping from 
interval $[0,T]$ into $\CS$: $\xx_i(0) = q_0$, $\xx_i(T) = q_{n+1}$ for all $i$. As illustrated in (Figure~\ref{optim_grad}), let 
us denote by $\tcolli$ the abscissa of the first collision detected on path 
$\xx_{i+1}$, which previous iteration $\xx_i$ was collision-free. Thus in 
configuration $\xx_{i+1}(\tcolli)$ a collision has been 
detected. Two cases are possible:
\begin{enumerate}
\item the collision occurred between two bodies of the robot: $\body_1$ and $
\body_2$, or
\item the collision occurred between a body of the robot $\body_1$ and the 
environment.
\end{enumerate}
In the rest of this section, we will only consider the first case. Reasoning 
about the second case is similar, except that the 
constraint is on the position of $\body_1$ with respect to the environment.

\begin{figure}[t]
	\centering
	\includegraphics[width=15.8cm]{contact-points.pdf}
	\caption{Bodies representation in collision-configuration (left) and 
	backtracked free-configuration (right). A contact point (i.e. any point in 
	the intersection of the bodies) $\po_c$ can be returned by the Flexible 
	Collision Library~\cite{fcl}, used to detect collisions. The red $(\perp)$ 
	hyperplan is orthogonal to ($\po_2$,$\mathbf{u}$) and represents the motion 
	direction which is allowed by the constraint.}
	\label{contact-points}
\end{figure}

The principle of the method is to compute a constraint on the backtracked free-
configuration $\xx_{i}(\tcolli)$ to avoid the encountered collision $\xx_{i+1}
(\tcolli)$. To handle this, we introduce a one-dimensional constraint based on the 
orthogonal direction of the encountered collision.

In the collision-configuration $\xx_{i+1}(\tcolli)$, let $\po_c\in \real^3$ be a 
contact point expressed in the global frame ${\cal R}_0$ 
(Figure~\ref{contact-points} left). We compute the respective positions $\po_1$ and 
$\po_2$ of $\po_c$ in 
the local frames of $\body_1$ and $\body_2$ (Figure~\ref{contact-points} right).

@@

Notice that $\mathbf{u}$ is well defined since configuration 
$\xx_{i}(t_{coll\ i})$ is collision-free.
Let $f$ be the function defined from $\CS^{wp}$ to $\real$ by:
$$
f (\xx) = \left(M_{1}^{-1} (\xx) M_2 (\xx) \po_2 - \po_1 | \mathbf{u}\right)
$$
The constraint is defined for any path $\xx$ by:
\begin {equation}\label{eq:new-constraint}
f(\xx) - f(\xx_{i}) = 0.
\end {equation}
As this constraint is non-linear, it is linearized around $\xx_{i}$. Let $\conf_{k\,i}$ denote the waypoint $k$ along path $\xx_i$.
The constraint on the position at $t_{coll\,i}$ is a function that depends only
on two waypoints $\conf_{k\,i}$, $\conf_{k+1\,i}$, with $0\leq k\leq n$. There exists 
$\beta\in[0,1]$ such that for any $j$:
$$
\xx_j (t_{coll\,i}) = \conf_{k\,j} + \beta (\conf_{k+1\,j} - \conf_{k\,j})
$$
The path Jacobian $\frac{\partial f}{\partial \xx}(\xx_i)$ is thus built by matrix 
blocks using the previous constraint Jacobian $\frac{\partial f}{\partial \conf}$ 
and $\beta$, corresponding to \texttt{computeCollisionConstraint} in 
(Algorithm~\ref{algo:gradient}).
Then, a line is added in the contraint Jacobian matrix (function 
\texttt{addCollisionConstraint}), and the constraint 
Jacobian becomes:
\begin {align*}
\Jf_{i+1} &= \left(\begin{array}{c}L_1 \\ \vdots \\ L_{k+1}\end{array}\right)\ \mbox {with}\\
L_{k+1} &= \frac{\partial f}{\partial \xx}(\xx_i)
\end{align*}

@@

Then we refer to~\cite{nocedal2006numerical} for solving constrained quadratic 
programs (QP) (function \texttt{computeIterate} in 
(Algorithm~\ref{algo:gradient})).

@@

\subsection*{Convergence Analysis}
For iteratively constrained problem, the risk of falling in an infinite loop 
that keeps adding a redundant constraint exists.
In this section, we demonstrate that, given the constraint definition and the 
function \texttt{solveRedundantConstraint}, our 
algorithm will converge.
Under the previous notations, it is straight-forward that
\begin{equation}\label{eq:decreasing-kernel}
\kernel \Jf_{i+1} \subset \kernel \Jf_{i}
\end{equation}
Let us assume that:
\begin{equation}\label{eq:asumption-1}
L_{k+1}\p_i \not= 0.
\end{equation}
We will elaborate later on this asumption. This means that $\p_{i}\notin\kernel J_{i+1}$, and as $\p_{i}\in\kernel J_{i}$, then:
$$
\kernel J_i \not= \kernel J_{i+1}.
$$
From~(\ref{eq:decreasing-kernel}), we deduce that:
$$
\dim (\kernel J_{i+1}) < \dim (\kernel J_i).
$$
This results proves that each added constraint is linearly independant from the 
previous ones, and thus our algorithm terminates in a finite number of iterations.

\subsubsection*{About asumption~(\ref{eq:asumption-1})}%
Let us consider the following expression:
$$
\frac{\partial f}{\partial \xx}(\xx_{i+1})\p_i
$$
This expression represents the velocity of point $\po_2$ in the reference frame of $\body_1$ projected onto $\mathbf{u}$ along the motion of the robot defined by the following trajectory:
\begin{equation}\label{eq:trajectory}
\conf (t) = \xx_{i}(t_{coll\ i}) + t\, \alpha_i \p_{i}.
\end{equation}
The motion is defined by taking a constant abscissa $t_{coll\ i}$ and be moving the 
path along iterate $\p_i$. Note that configuration $\xx_{i+1}(t_{coll\ i})$ is 
reached for $t=1$.
This expression being equal to zero requires either
\begin {enumerate}
\item that $\po_1$ and $\po_2$ come to contact at zero velocity, or
\item the velocity of $\po_2$ in the reference frame of $\body_1$ is orthogonal to 
$\mathbf{u}$ at contact ($t=1$).
\end {enumerate}
The first case is very unlikely and almost never encountered in real applications, 
except if the user builds such a case on purpose.
The second case may appear and can be noted by a rank loss of $\Jf_{i+1}$. In this 
case, we do not directly insert the new constraint. Instead, we process intermediate 
iterations until the redundancy vanishes, as explained in function 
\texttt{solveRedundantConstraint}.

>>>>

\title{A simple path optimization method for motion planning$^{1}$ \thanks{$^{1}$Note for the reviewers: This paper is an extended version of a paper submitted to IEEE ICRA 2016~\cite{pathOptimICRA_HAL}. The current version includes a problem formulation, a convergence analysis and many experimental results (including industrial examples) which are absent from the conference version.}}

\author{Myl\`{e}ne Campana$^{\ast}$ \thanks{$^\ast$Corresponding author. Email: \href{mailto:mcampana@laas.fr}{mcampana@laas.fr}}, Florent Lamiraux and Jean-Paul Laumond
\\\vspace{6pt}
{\em{LAAS-CNRS, 7 av. Colonel Roche, 31400 Toulouse, FRANCE}}
}
\maketitle

@@

\section{Introduction}
Motion planning for systems in cluttered environments has been addressed for more
than thirty years~\cite{ref-motionplan}. 
To explore the connected components of collision-free configuration spaces, pionnering contributions in the 90’s introduced certain levels of random searches, i.e. random walks in~\cite{potentielBarraquandLatombe}, random sampling in 
~\cite{KavrakiLatombePRM} and~\cite{LaValleKuffnerRRT}. Today most motion planners are inspired by these seminal approaches.
The main issue using these techniques is that the computed path makes unnecessary 
detours and needs to be post-processed before being executed by a virtual or real 
robot. Alternative strategies exist however.
\begin{itemize}
\item Planning by path-optimization~\cite{itomp2012},~\cite{voronoiOMP} where
obstacle avoidance is handled by constraints or cost using computation of the
nearest obstacle distance. Most of these planners are using non-linear
optimization~\cite{BettsNonlinopt} under constraints. Such planners provide close-to-optimality paths and have smaller time computation
for easy problems, but they are mostly unable to solve narrow passage issues.

\item Optimal random sampling~\cite{KaramanPRMstarRRTstar} are also close to an
optimal solution, but computation time is significantly higher than classical
approaches.
%Moreover the roadmap is only valid for a given motion planning problem. 
\end{itemize}

@@

For this shortening purpose, random shortcut (RS) methods are
still very 
popular~\cite{Sekhavat-Svestka1998,Geraerts04clearancebased,HauserFastSmooth}. 
However, RS requires fine 
tuning of the termination condition and is no efficient for long 
trajectories where only a small part needs to be optimized, 
as in the (Figure~\ref{local_box_optim}). 
The (Figure~\ref{decoupled_DOF_optimization}) presents another situation where RS 
will always fail to optimize the initial path, since it cannot decouple the 
robot degrees of freedom (DOF) on which the optimization occurs. This problem is 
addressed by our method.

\begin{figure}
	\centering
	\includegraphics[width=7.5cm]{decoupled_DOF_optimization.pdf}
	\caption{Example of a path in $\real^3$ which random shortcut will never manage to 
	optimize: each shortcut tentative will provide a collision or will not 
	decrease the path length. The optimal path belongs to the $x-y$ plane 
	containing $\conf_{init}$ and $\conf_{final}$.}
	\label{decoupled_DOF_optimization}
\end{figure}

On the other hand, numerical
optimization methods like CHOMP~\cite{chompIjrr} can be used as a
post-processing step. They have clear termination conditions, but collision
avoidance is handled by inequality constraints sampled at many points along
the trajectory. These methods therefore require a pre-processing step of the
robot (and/or environment) model in order to make 
it simpler:~\cite{chompIjrr} covers PR2 bodies with spheres, 
while~\cite{convexOptimMotplan} needs to decompose objects into convex subsets.

Finally it should be noticed that optimality in robot motion is a notion that should be clarified. Most of the time motion planners provide an optimized motion, which is not optimal at all, but is the output of a given optimization method. When optimal motions exist, numerical
algorithms mostly fail in accounting for their combinatorial structure. Moreover, optimization algorithms bypass (not overcome) the question of the existence of optimal motions~\cite{LaumondOptim}. In that perspective, a path optimization algorithm has to be evaluated with respect to other existing optimization techniques, from qualitative properties and from computational performance.

@@

\section{RELATED WORK}
Previous work on path optimization has been conducted. CHOMP algorithm~\cite{chompIjrr} optimizes an initial guess provided as
input. It minimizes a time invariant cost function using efficient covariant
hamiltonian gradient descent. The cost is quantified by non-smooth parts (with
high velocities) and an obstacle avoidance term, provided by the distance to the 
nearest obstacle for each iteration of the trajectory. Calculating these nearest 
distances however is time-consuming because the distances between all pairs of 
objects must be computed at each time step along the path. To reduce the 
computation time, the method starts by building offline a map of distances that 
will be called during the optimization at the requested time. Besides, meshes 
are pre-processed into bounding spheres so that distances are computed faster 
at the cost of a geometry approximation.

@@

Some heuristics use random shortcuts on the initial guess combined with a 
trajectory re-building that returns ${C}^1$ shortcuts made of parabolas and lines 
(bang-bang control)~\cite{HauserFastSmooth}. These local refined trajectories 
are time-optimal since they comply with acceleration and velocity constraints. 
~\cite{Guernane2011}, also based on random shooting, is guiding 
configuration-generation with local holonomic considerations. 
But this method remains only 
locally optimal, and is not addressing high-DOF problems.
~\cite{Geraerts04clearancebased}, whereas just implemented for freeflyer robots, is 
using medial axis retractation for clearance and a PRS (applied only on random DOF) 
which can address the problem of (Figure~\ref{decoupled_DOF_optimization}) but is 
relatively slower than RS. Moreover, PRS is not using the information returned by 
the checker of which links are colliding to guide which DOF or group of DOF are 
relevant to shortcut, contrary to our method.

In some way, our method shares similarities with~\cite{PanSmoothSplineShort} 
since this latter work relies on collision checking and backtracks when an 
iteration is detected in collision, instead of trying to constantly satisfy 
distance constraints. Unlike our method however, collision constraints are 
handled by interpolating configurations which, at some points of the trajectory, 
freeze the whole robot configuration instead of a relevant subpart. Thus, this 
method cannot solve the (Figure~\ref{decoupled_DOF_optimization}) problem.

@@

A robot is defined by a kinematic chain composed of a tree of joints. We denote 
by $(J_1,\cdots,J_m)$ the ordered list of joints. Each joint $J_i$, $1\leq i\leq 
m$ is represented by a mapping from a sub-manifold of $\real^{n_i}$, where $n_i$ 
is the dimension of $J_i$ in the configuration space, to the space 
of rigid-body motions $SE(3)$. The rigid-body motion is the position of the joint 
in the frame of its parent. In the examples shown in this paper, we consider four 
types of joints described in (Table~\ref{tab:joints}).
\begin{table}
\centerline {
  \begin{tabular}{cccc}
    Name & dimension & config space & velocity\\
    \hline
    translation & 1 & $\real$ & $\real$\\
    bounded rotation & 1& $\real$ & $\real$\\
    unbounded rotation & 2 & $\Sone\subset\real^2$ & $\real$\\
    $SO(3)$ & 4 & $\Sthree\subset\real^4$ & $\real^3$
  \end{tabular}
}
\caption {Translation and rotation joint position are defined by one 
parameter corresponding respectively to the translation along an axis and a 
rotation angle around an axis. Unbounded rotation is defined by a point on the 
unit circle of the plane: two parameters corresponding to the cosine and the 
sine of the rotation angle. $SO(3)$ is defined by a unit quaternion. The 
velocity of translation and unbounded rotation joints is the derivative of the 
configuration variable. The velocity of an unbounded rotation joint corresponds 
to the angular velocity. The velocity of a $SO(3)$ joint is defined by the 
angular velocity vector $\omega\in\real^3$.}
\label{tab:joints}
\end{table}
A configuration of the robot
$$\conf = (\underbrace{q_1,\cdots,q_{n_1}}_{J_1},\underbrace{q_{n_1+1},
\cdots,q_{n_1+n_2}}_{J_2},\cdots q_n),\ n\triangleq\sum_{i=1}^m n_i$$
is defined by the concatenation of the joint configurations. The configuration 
space of the robot is denoted by $\CS\subset\real^n$.

@@

\subsubsection{Operations on configurations and vectors}%
By analogy with the case 
where the configuration space is a vector space, we define the following 
operators between configurations and vectors:
$$
\conf_2 - \conf_1 \in \real^p, \ \conf_1, \conf_2\in\CS
$$
is the constant velocity moving from $\conf_1$ to $\conf_2$ in unit time, and
$$
\conf + \dot{\conf}\in\CS, \ \conf\in\CS \ \dot{\conf}\in\real^p
$$
is the configuration reached from $\conf$ after following constant velocity $
\dot{\conf}$ during unit time.

@@

\subsubsection{Cost}

let $W\in\real^{p\times p}$ be a diagonal matrix of weights:
$$
W=\left(\begin{array}{cccccccccc}
w_1 I_{p_1}       &        &  0  \\
    & w_2 I_{p_2} &        &   \\
    &            & \ddots &   \\
  0 &            &        & w_m I_{p_m}
\end{array}\right)
$$
where $I_{p_i}$ is the identity matrix of size $p_i$ and $w_i$ is the weight 
associated to the joint $J_i$. We define the length of the straight interpolation 
between two configurations as:
$$
\|\conf_2 - \conf_1\|_{W} \triangleq \sqrt{(\conf_2 - \conf_1)^T W^2 (\conf_2 - 
\conf_1)}.
$$
Weights are used to homogenize translations and rotations in the velocity vector. 
For rotations, the weight is equal to the maximal distance of the robot bodies 
moved by the joint to the center of the joint, in a given configuration.

Given $\conf_0$ and $\conf_{wp+1}$ fixed, the cost we want to minimize is defined 
by
$$
\cost (\xx) \triangleq \frac{1}{2}\sum_{i=1}^{wp+1} \lambda_{i-1} \|\conf_{i}-\conf_{i-1}\|_{W}
^{2}
$$
where the $\lambda_{i-1}$ coefficients will be explained in the results 
section. For now it is assumed that $(\lambda_{i-1})_{1\leq i\leq wp+1}=1$.

@@

The gradient of the cost function $\nabla \cost (\xx)$ is computed as follows:
$$
\nabla \cost (\xx) = 
\left( (\lambda_{i}(\conf_{i+1} - \conf_{i})^T - \lambda_{i+1}(\conf_{i+2} - \conf_{i+1})^T) \weight^2 \right)_{i=0\cdots wp-1}
$$

Let $\mbox{H}$ denote the constant Hessian of the cost function:
$$
\mbox{H} = \left(\begin{array}{cccccc}
(\lambda_{0}+\lambda_{1})\weight^2 & -\lambda_{1}\weight^2 & 0 & \cdots & & 0 \\
-\lambda_{1}\weight^2 & (\lambda_{1}+\lambda_{2})\weight^2 & -\lambda_{2}\weight^2 & 0 & \cdots & 0 \\
0 & -\lambda_{2}\weight^2 &  (\lambda_{2}+\lambda_{3})\weight^2 & -\lambda_{3}\weight^2 & 0 & \vdots \\
\vdots & \ddots & \ddots & \ddots & \ddots & \vdots\\
\vdots & & \ddots & \ddots & \ddots & \vdots\\
0 & \cdots  & 0 & -\lambda_{wp-2}\weight^2 & (\lambda_{wp-2}+\lambda_{wp-1})\weight^2 & -\lambda_{wp-1}\weight^2 \\
0 & \cdots &  \cdots & 0 & -\lambda_{wp-1}\weight^2 & (\lambda_{wp-1}+\lambda_{wp})\weight^2  \\
\end{array}\right)
$$

\subsection {Resolution}
We assume that the direct interpolation between the initial and final configurations contains collisions. An iteration is described as follow:
\begin{equation}\label{eq:iteration-1}
\begin{split}
& \pii =  -H^{-1} \nabla \cost(\xx_i)^{T} \\
& \xx_{i+1} =  \xx_{i} + \alpha_i \pii
\end{split} 
\end{equation}
where $\alpha_i$ is a real valued parameter. Taking $\alpha_i=1$ yields the 
unconstrained minimal cost path, i.e. all waypoints aligned on the straight line 
between $q_0$ and $q_{wp+1}$. Since this solution is in collision, we set 
$\alpha_i = \alpha$ 
where $\alpha$ is a parameter that will be explained in the algorithm section.

@@

\begin{figure}
	\centering
	\includegraphics[width=9cm]{optim_grad.pdf}
	\caption{Illustration of one iteration of our optimization. $\xx_{i+1}$ 
	appears 
	to be in collision with the obstacle, the first colliding configuration 
	$\xx_{i+1}(\tcolli)$ is returned by the continuous collision checker. The 
	corresponding constraint will be computed in the backtracked configuration 
	$\xx_{i}(\tcolli)$.}
	\label{optim_grad}
\end{figure}

@@

\subsubsection{New constraint}

Each path $\xx_i$ is a mapping from 
interval $[0,1]$ into $\CS$: $\xx_i(0) = q_0$, $\xx_i(1) = q_{wp+1}$ for all $i$. As illustrated in (Figure~\ref{optim_grad}), let 
us denote by $\tcolli$ the abscissa of the first collision detected on path 
$\xx_{i+1}$, which previous iteration $\xx_i$ was collision-free. Thus in 
configuration $\xx_{i+1}(\tcolli)$ a collision has been 
detected. Two cases are possible:
\begin{enumerate}
\item the collision occurred between two bodies of the robot: $\body_1$ and $
\body_2$, or
\item the collision occurred between a body of the robot $\body_1$ and the 
environment.
\end{enumerate}
In the rest of this section, the first case will only be considered. Reasoning 
about the second case is similar, except that the 
constraint is on the position of $\body_1$ with respect to the environment.

\begin{figure}
	\centering
	\includegraphics[width=15.8cm]{contact-points.pdf}
	\caption{Bodies representation in collision-configuration (left) and 
	backtracked free-configuration (right). A contact point (i.e. any point in 
	the intersection of the bodies) $\po_c$ can be returned by the Flexible 
	Collision Library~\cite{fcl}, used to detect collisions. The red $(\perp)$ 
	hyperplan is orthogonal to ($\po_2$,$\mathbf{u}$) and represents the motion 
	direction which is allowed by the constraint.}
	\label{contact-points}
\end{figure}

The principle of the method is to compute a constraint on the backtracked 
free-configuration $\xx_{i}(\tcolli)$ to avoid the encountered collision $\xx_{i+1}
(\tcolli)$. To handle this, we introduce a one-dimensional constraint based on the 
orthogonal direction of the encountered collision.

In the collision-configuration $\xx_{i+1}(\tcolli)$, let $\po_c\in \real^3$ be a 
contact point expressed in the global frame 
(Figure~\ref{contact-points} left). We compute the respective positions $\po_1$ and 
$\po_2$ of $\po_c$ in 
the local frames of $\body_1$ and $\body_2$ (Figure~\ref{contact-points} right).

@@

Notice that $\mathbf{u}$ is well defined since configuration 
$\xx_{i}(t_{coll\ i})$ is collision-free.
Let $f$ be the function defined from $\CS^{wp}$ to $\real$ by:
$$
f (\xx) = \left(M_{1}^{-1} (\xx) M_2 (\xx) \po_2 - \po_1 | \mathbf{u}\right)
$$
The constraint is defined for any path $\xx$ by:
\begin {equation}\label{eq:new-constraint}
f(\xx) - f(\xx_{i}) = 0.
\end {equation}
As this constraint is non-linear, it is linearized around $\xx_{i}$. This step is described in a further section. Then, a line is added in the contraint Jacobian matrix (function 
\texttt{addCollisionConstraint}), and the constraint 
Jacobian becomes:
\begin {align*}
\Jf_{i+1} &= \left(\begin{array}{c}L_1 \\ \vdots \\ L_{k+1}\end{array}\right)\ \mbox {with}\\
L_{k+1} &= \frac{\partial f}{\partial \xx}(\xx_i)
\end{align*}

@@

Finally, we refer to~\cite{nocedal2006numerical} for solving constrained quadratic 
programs (QP) (function \texttt{computeIterate} in 
(Algorithm~\ref{algo:gradient})).

@@

\vspace{0.2cm}

\subsubsection{Linearized constraint computation} \label{sec:lin_constr_compt}

Let $\conf_{k\,i}$ denote the waypoint $k$ along path $\xx_i$.
The constraint on the position at $t_{coll\,i}$ is a function that depends only
on two waypoints $\conf_{k\,i}$, $\conf_{k+1\,i}$, with $0\leq k\leq wp$.
There exists $\beta\in[0,1]$ such that for any $j$:
$$
\xx_j (\tcolli) = \conf_{k\,j} + \beta (\conf_{k+1\,j} - \conf_{k\,j})
$$
The path Jacobian $\frac{\partial f}{\partial \xx}(\xx_i)$ is thus built by matrix 
blocks using the previous constraint Jacobian $\frac{\partial f}{\partial \conf}$ 
and $\beta$, corresponding to \texttt{computeCollisionConstraint} in 
(Algorithm~\ref{algo:gradient}).

\subsection{Convergence Analysis}
For iterative algorithms, the risk of falling in an infinite loop 
that keeps adding a redundant constraint exists.
In this section, we give some asumptions under which our 
algorithm will converge, given the constraint definition and the 
function \texttt{solveRedundantConstraint}.
From the definition of $\Jf_{i}$, it is straightforward that:
\begin{equation}\label{eq:decreasing-kernel}
\kernel \Jf_{i+1} \subset \kernel \Jf_{i}
\end{equation}
Let us assume that:
\begin{equation}\label{eq:asumption-1}
L_{k+1}\p_i \not= 0.
\end{equation}
We will elaborate later on this asumption. This means that $\p_{i}\notin\kernel J_{i+1}$, and as $\p_{i}\in\kernel J_{i}$, then:
$$
\kernel J_i \not= \kernel J_{i+1}.
$$
From~(\ref{eq:decreasing-kernel}), we deduce that:
$$
\dim (\kernel J_{i+1}) < \dim (\kernel J_i).
$$
This result proves that under Asumption~(\ref{eq:asumption-1}), each added constraint 
is linearly independent from the 
previous ones, and thus our algorithm terminates in a finite number of iterations.

\vspace{0.2cm}
